Parameter estimation

### 2023-04-22 SFE-article/Sections/Gouverning_equation_derivation.tex
21.04

	Following the work of \citet{Anderson1995}, the governing equations for compressible fluid with non-uniform cross-section can be obtained. Let's assume that any properties of the flow are uniform across any given cross-section of an extractor. The variation of the cross-section might be an result of partial filling of an extractor or its irregular shape. In reality, such a flow is two-dimensional, because with the area changing as a function of $z$, in actuality there will be flow-field variations in both directions. The assumption of quasi-one-dimensional flow dictates that the flow properties are function of $z$ only. The equations described by quasi-one-dimensional assumption hold: (1) mass conservation, (2) Newton's second law, and (3) energy conservation. To ensure that these physical principles are satisfied the modified governing equation can be derived.	Let's start with the integral form of the continuity equation:

@@

	\begin{figure}[h]
		\centering
		\begin{tikzpicture}
			\draw [fill=blue!20!white,opacity=0.5] (2.25,2.44) -- (2.25,0.57) -- (1.25,0.31) -- (1.25,2.69) -- cycle;%smmall triangle
			\draw[dashed,color=gray] (0,0) arc (-90:90:0.5 and 1.5);% right half of the left ellipse
			\draw[semithick] (0,0) -- (4,1);% bottom line
			\draw[semithick] (0,3) -- (4,2);% top line
			\draw[semithick] (0,0) arc (270:90:0.5 and 1.5);% left half of the left ellipse
			\draw[semithick] (4,1.5) ellipse (0.166 and 0.5);% right ellipse
			\draw (-1,1.5) node {$A_1$};
			\draw (3.3,1.5) node {$A_2$};
			\draw[|-,semithick] (0,-1.5) -- (4,-1.5);
			\draw[|->,semithick] (4,-1.5) -- (4.5,-1.5);
			\draw (0,-1) node {$z=0$};
			\draw (4,-1) node {$z=L$};
			\draw (1.75,-0.25) node{\scriptsize $ dz $}; %a number
			\draw[|<->|] (1.25,-0.5) -- (2.25,-0.5); %lenght indicator
			\draw[dashed] (1.25,-0.5) -- (1.25,0.3); %dashed line
			\draw[dashed] (2.25,-0.5) -- (2.25,0.58); %dashed line
		\end{tikzpicture}
		\caption{Control volume for deriving partial differential equation for unsteady, quasi-one-dimensional flow}
		\label{fig: control_volume}
	\end{figure}

	This control volume is a slice of an extractor, where the infinitesimal thickens of the slice is $dz$. On the left side of the control volume, consistent with the quasi-one-dimensional assumptions, the density, velocity, pressure and internal energy denoted by $\rho_f$, $V$, $P$, and $e$, respectively, are uniform over the are $A$. Similarly, on the right side of the control volume, the density, velocity, pressure, and internal energy $\rho_f+d\rho_f$, $V+dV$, $P+dP$, and $e+de$, respectively, are uniform over the area available for fluid phase $A_f+dA_f$. Applied to the control volume in Fig. \ref{fig: control_volume}, the volume integral in Eq. \ref{EQ: Continuity_integral_general} becomes, in the limit as $dz$ becomes very small,

@@

	where the minus sign on the leading term on the right-hand side is due to the vectors $\textbf{V}$ and $\textbf{dS}$ pointing in opposite directions over the left of the control volume, and hence the dot product is negative. Expending the triple product term

@@

	In the limit as $dz$ becomes very small, the terms involving products of the differential in Eq. \ref{EQ: Triple_Prod_Cont_Surf}, sych as $\rho_f dV dA_f,~d\rho_f dV dA_f$, go to zero much faster than those terms involving only one differential. Hence, all terms involving products of differentials can be dropped, yielding in the limit as $dz$ becomes very small

@@

	Above partial differential equation form of the continuity equation suitable for unsteady, quasi-one-dimensional flow. It ensures that mass is conserved for this mode of the flow. The $A_f(z)$ is an arbitrary function, which describe change of the cross-section of an extractor. The function $A_f(z)$ can be defined as $A_f(z) = \textbf{A} \phi(z)$, where $\phi$ is the bed porosity and $\textbf{A}$ is the cross-section of an empty extractor.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \textbf{A} \phi (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f \textbf{A} \phi (z) V \right)}{\partial z} = 0
		\end{equation}
	}
	The equation can be simplified by cancel out a constant $\textbf{A}$

@@

	If so called superficial velocity is defined $u=\phi V$, the mass continuity becomes

@@

	The transport of a chemical species, in this case a solute, can be described by analogous equation to the Eq. \ref{EQ: Continuity_integral_general} with additional terms on the right-hand side. The first term on the right-hand side describes that a substance goes from high density regions to low density regions and is based on the Fick's law $\left( J_{diff} = D_e^M\frac{\partial c_f}{\partial z} \right)$. The other term correspond for the mass transfer between solid and fluid phases, which is treated as a source term.

@@

	The surface integrals in the limit of $dz$ becomes

@@

	The equation can be further simplified if $\frac{\partial u}{\partial z} = \frac{\partial \phi}{\partial z} = D_e^M = 0$, which corresponds to the assumptions of constant velocity along the bed( which might be a case of isothermal and low-mach number flow), constant porosity( which comes from the assumption of constant area for both solid and fluid phase) and no radial diffusion.

@@

	Similarly to the mass conservation, the momentum conservation is derived for inviscid fluid with no body forces

@@

	The evaluation of the pressure force term on the right side of Eq. \ref{EQ: Momentum_Integral_Form} can be understood based on the Fig. \ref{fig: Forces_Momentum_Control_Volume}. Here, the $z$ components of the vector $PdS$ are shown on all four side of the control volume. Remember that $\textbf{dS}$ is assumed to points away from the control volume; hence  any $z$ component $\left( PdS \right)_z$ that acts toward the left (in the negative $z$ direction) is a negative quantity, and any $z$ component that acts toward the right (in the positive $z$ direction) is a positive quantity. Also note that the $z$ component of $P\textbf{dS}$ acting on the top and the bottom inclined faces of the control volume in Fig. \ref{fig: Forces_Momentum_Control_Volume} can be expressed as the pressure $P$ acting on the component of the inclined are projected perpendicular to the $z$ direction, $dA_f/2$; hence, the contribution of each inclined face (top or bottom) to the pressure integral in Eq. \ref{EQ: Momentum_Integral_Form} is $-P(dA_f/2)$. All together, the right-hand side of Eq. \ref{EQ: Momentum_Integral_Form} is expressed as follows:

@@

	Cancelling like terms and ignoring products of differentials, equation above becomes in the limit $dz$ becoming  very small

@@

	Dividing above equation by $dz$ and taking the limit as $dz$ goes to zero, we obtain

@@

	Expanding the derivatives on the left-hand size of above equation and cancelling like terms, gives

@@

	Dividing above equation by $A_f$ the non-conservative form of the momentum can be obtained

@@

	By expanding all the terms of equation above, we get

@@

	The Eq. \ref{EQ: Non-conservative_Momentum_Const_Porosity} represents non-conservative form of the momentum equation for quasi-one-dimensional flow with no body forces and constant porosity.

@@

	Neglecting products of differential and cancelling like terms, the above equation becomes

@@

	Expanding right-hand side and dividing by $A_f$, the above equation becomes

@@

	Equation \ref{EQ: Energy_equation_nonconservative} is the non-conservative for of the energy equation expressed in therms of internal energy, appropriate to unsteady quasi-one-dimensional flow. The reason for obtaining the energy equation in the form of Eq. \ref{EQ: Energy_equation_nonconservative} is that, for a calorically perfect gas, it leads directly to a form of the energy equation in therms of temperature $T$. For calorically perfect gas ${\footnotesize e_f=C_vT}$

@@

%	{TODO: \color{red}finish derivation of energy equation for real gas and introduce homogenous version of the energy equation}

>>>>

	Following the work of \citet{Anderson1995}, the governing equations for compressible fluid with non-uniform cross-section can be obtained. Let's assume that any properties of the flow are uniform across any given cross-section of an extractor. The variation of the cross-section might result from the partial filling of an extractor or its irregular shape. In reality, such a flow is two-dimensional because the area changes as a function of $z$, and there is a flow-field variation in both directions. The assumption of quasi-one-dimensional flow dictates that the flow properties are a function of $z$ only. The equations described by quasi-one-dimensional assumption hold: (1) mass conservation, (2) Newton's second law, and (3) energy conservation. The modified governing equation can be derived to ensure that these physical principles are satisfied.	Let's start with the integral form of the continuity equation:

@@

	\begin{figure}[h]
		\centering
		\begin{tikzpicture}
			\draw [fill=blue!20!white,opacity=0.5] (2.25,2.44) -- (2.25,0.57) -- (1.25,0.31) -- (1.25,2.69) -- cycle;%smmall triangle
			\draw[dashed,color=gray] (0,0) arc (-90:90:0.5 and 1.5);% right half of the left ellipse
			\draw[semithick] (0,0) -- (4,1);% bottom line
			\draw[semithick] (0,3) -- (4,2);% top line
			\draw[semithick] (0,0) arc (270:90:0.5 and 1.5);% left half of the left ellipse
			\draw[semithick] (4,1.5) ellipse (0.166 and 0.5);% right ellipse
			\draw (-1,1.5) node {$A_1$};
			\draw (3.3,1.5) node {$A_2$};
			\draw[|-,semithick] (0,-1.5) -- (4,-1.5);
			\draw[|->,semithick] (4,-1.5) -- (4.5,-1.5);
			\draw (0,-1) node {$z=0$};
			\draw (4,-1) node {$z=L$};
			\draw (1.75,-0.25) node{\scriptsize $ dz $}; %a number
			\draw[|<->|] (1.25,-0.5) -- (2.25,-0.5); %lenght indicator
			\draw[dashed] (1.25,-0.5) -- (1.25,0.3); %dashed line
			\draw[dashed] (2.25,-0.5) -- (2.25,0.58); %dashed line
		\end{tikzpicture}
		\caption{Control volume for deriving the partial differential equation for unsteady, quasi-one-dimensional flow}
		\label{fig: control_volume}
	\end{figure}

	This control volume is a slice of an extractor, where the infinitesimal thickness of the slice is $dz$. On the left side of the control volume, consistent with the quasi-one-dimensional assumptions, the density, velocity, pressure, and internal energy denoted by $\rho_f$, $V$, $P$, and $e$, respectively, are uniform over the area $A$. Similarly, on the right side of the control volume, the density, velocity, pressure, and internal energy $\rho_f+d\rho_f$, $V+dV$, $P+dP$, and $e+de$, respectively, are uniform over the area available for fluid phase $A_f+dA_f$. Applied to the control volume in Fig. \ref{fig: control_volume}, the volume integral in Eq. \ref{EQ: Continuity_integral_general} becomes, in the limit as $dz$ becomes very small,

@@

	The minus sign on the leading term on the right-hand side is due to the vectors $\textbf{V}$ and $\textbf{dS}$ pointing in opposite directions over the left of the control volume, and hence the dot product is negative. Expanding the triple product term

@@

	In the limit as $dz$ becomes very small, the terms involving products of the differential in Eq. \ref{EQ: Triple_Prod_Cont_Surf}, such as $\rho_f dV dA_f,~d\rho_f dV dA_f$, go to zero much faster than those terms involving only one differential. Hence, all terms involving products of differentials can be dropped, yielding in the limit as $dz$ becomes very small

@@

	The above partial differential equation form of the continuity equation is suitable for unsteady, quasi-one-dimensional flow. It ensures that mass is conserved for this mode of flow. The $A_f(z)$ is an arbitrary function that describes a change in the cross-section of an extractor. The function $A_f(z)$ can be defined as $A_f(z) = \textbf{A} \phi(z)$, where $\phi$ is the bed porosity and $\textbf{A}$ is the cross-section of an empty extractor.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \textbf{A} \phi (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f \textbf{A} \phi (z) V \right)}{\partial z} = 0
		\end{equation}
	}
	The equation can be simplified by canceling out a constant $\textbf{A}$

@@

	If so-called superficial velocity is defined as $u=\phi V$, the mass continuity becomes

@@

	The transport of a chemical species, in this case, a solute, can be described by an analogous equation to the Eq. \ref{EQ: Continuity_integral_general} with additional terms on the right-hand side. The first term on the right-hand side describes that a substance goes from high-density regions to low-density regions and is based on the Fick's law $\left( J_{diff} = D_e^M\frac{\partial c_f}{\partial z} \right)$. The other term corresponds to the mass transfer between solid and fluid phases, which is treated as a source term.

@@

	The surface integrals in the limit of $dz$ become

@@

	The equation can be further simplified if $\frac{\partial u}{\partial z} = \frac{\partial \phi}{\partial z} = D_e^M = 0$, which corresponds to the assumptions of constant velocity along the bed( which might be a case of isothermal and low-Mach number flow), constant porosity( which comes from the assumption of constant area for both solid and fluid phase) and no radial diffusion.

@@

	Similarly to mass conservation, momentum conservation is derived for inviscid fluid with no body forces

@@

	The evaluation of the pressure force term on the right side of Eq. \ref{EQ: Momentum_Integral_Form} can be understood based on the Fig. \ref{fig: Forces_Momentum_Control_Volume}. Here, the $z$ components of the vector $PdS$ are shown on all four sides of the control volume. Remember that $\textbf{dS}$ is assumed to point away from the control volume; hence  any $z$ component $\left( PdS \right)_z$ that acts toward the left (in the negative $z$ direction) is a negative quantity. Any $z$ component that acts toward the right (in the positive $z$ direction) is a positive quantity. Also note that the $z$ component of $P\textbf{dS}$ acting on the top and the bottom inclined faces of the control volume in Fig. \ref{fig: Forces_Momentum_Control_Volume} can be expressed as the pressure $P$ acting on the component of the inclined is projected perpendicular to the $z$ direction, $dA_f/2$; hence, the contribution of each inclined face (top or bottom) to the pressure integral in Eq. \ref{EQ: Momentum_Integral_Form} is $-P(dA_f/2)$. All together, the right-hand side of Eq. \ref{EQ: Momentum_Integral_Form} is expressed as follows:

@@

	Canceling like terms and ignoring products of differentials, the equation above becomes in the limit $dz$ becoming  very small

@@

	Dividing the above equation by $dz$ and taking the limit as $dz$ goes to zero, we obtain

@@

	Expanding the derivatives on the left-hand side of the above equation and canceling like terms, gives

@@

	Dividing the above equation by $A_f$ the non-conservative form of the momentum can be obtained

@@

	By expanding all the terms of the equation above, we get

@@

	The Eq. \ref{EQ: Non-conservative_Momentum_Const_Porosity} represents the non-conservative form of the momentum equation for quasi-one-dimensional flow with no body forces and constant porosity.

@@

	Neglecting products of differential and canceling like terms, the above equation becomes

@@

	Expanding the right-hand side and dividing by $A_f$, the above equation becomes

@@

	Equation \ref{EQ: Energy_equation_nonconservative} is the non-conservative form of the energy equation expressed in terms of internal energy, appropriate to unsteady quasi-one-dimensional flow. The reason for obtaining the energy equation in the form of Eq. \ref{EQ: Energy_equation_nonconservative} is that, for a calorically perfect gas, it leads directly to a form of the energy equation in terms of temperature $T$. For calorically perfect gas ${\footnotesize e_f=C_vT}$

@@

%	{TODO: \color{red}finish derivation of energy equation for real gas and introduce a homogenous version of the energy equation}
